\emph{Light Exposure Behavior Assessment (LEBA)}: Develop of a novel instrument to capture light exposure-related behaviours

### 2021-08-31 manuscript.Rmd

```{r setup, include = FALSE}
library(papaja) # devtools::install_github("crsh/papaja")
library(readr)
library(readxl)
library(lavaan)
library(semPlot)
library(semTools)
library(MOTE)
library(psych)
library(dlookr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
r_refs("references.bib")
```

@@

We used R (version 4.1.0), including several R-packages for our analyses. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the  ordinal nature of our response data we used polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (pa) a factor extraction method with varimax rotation. PA is apparently robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by weighted least squares (WLS) extraction method as well. We used a combination factor indentification method including scree plot[@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method[@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers.  Additionally, to identify the simple structure we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018; @childEssentialsFactorAnalysis2006; @mulaikFoundationsFactorAnalysis2009; @watkinsStepbyStepGuideExploratory2020]

@@

qgraph(correlations$rho, cut =.30, details =T, posCol = "green", negCol = "orangered3", 
       labels = names(LB))
```

@@

```{r EFA, include=FALSE}
fa.5F.1 <- fa(r=correlations$rho, nfactors = 5, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)
A <- print(fa.5F.1, cut = .3, digits = 3, sort = TRUE)

@@

fa.5F.2 <- fa(r=correlations.red.5F.1$rho, nfactors = 5, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)
B <- print(fa.5F.2, cut = .3, digits = 3, sort = TRUE)

@@

C <- print(fa.5F.3, cut = .3, digits = 3, sort = TRUE)

@@

D <- print(fa.5F.4, cut = .3, digits = 3, sort = TRUE)

@@

```{r TabEFA5, results='asis'}
Fac.5 <- fa.5F.4 
Factorloading <- cbind(Fac.5$loadings, Fac.5$communality)
colnames(Factorloading)= c("F1", "F2","F3", "F4", "F5", "Communalities")
apa_table(unclass(Factorloading))
```

```{r communalities, include=FALSE}
min.com <- min(Fac.5$communality)
max.com <- max(Fac.5$communality)
```

The initial five-factor solution with all 48 items showed the presence of cross-loading items (item 42, 16, & 1) and poor factor loading (<.30) items (item 20,3 ,15, 17, 40, 4, 11, 39, 18, 45, 29, 25, 8 , & 46). At first we discard the items with poor  factor loading and ran another EFA on the remaining 34 items. This iteration of EFA also appeared as a misfit in terms of poor factor loading (Item 12, 22, 38, 6) and cross-loading (items 23, 31, 37, 48) . Another two rounds of EFA were conducted with gradually identifying problematic items and discarding them from the model. Finally, a five-factor EFA solution with 23 items was accepted with RMSR = 0.04, no loading smaller than .30 and no cross-loading greater than .30. The communalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. The histogram of the absolute values of non-redundant residual-correlations showed  correlations greater than the absolute value of .05 is ------[@desjardinsHandbookEducationalMeasurement2018a].
```{r FivRes, out.height= "50%",  fig.align='center', fig.cap=' Residulas of five-dactor solution'}
residual.5F =residuals(fa.5F.4, diag = FALSE, na.rm =T)
hist(residual.5F, col ="red", main = NULL, xlab = "Five factor oblique rotation: Residuals",ylim=c(0, 300))
abline(h =48, lwd = 2, lty = 2)
abline(v =.05,lwd = 2, lty = 2)
```

>>>>

```{r setup, include = FALSE}
library(papaja) # devtools::install_github("crsh/papaja")
library(readr)
library(readxl)
library(lavaan)
library(semPlot)
library(semTools)
library(MOTE)
library(psych)
library(dlookr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(correlation)
library(ggcorrplot)
r_refs("references.bib")
```

@@

We used R (version 4.1.0), including several R-packages for our analyses. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the  ordinal nature of our response data we used polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (pa) a factor extraction method with varimax rotation. PA is apparently robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by minimum residuals extraction method as well. We used a combination factor indentification method including scree plot[@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method[@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers.  Additionally, to identify the simple structure we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018; @childEssentialsFactorAnalysis2006; @mulaikFoundationsFactorAnalysis2009; @watkinsStepbyStepGuideExploratory2020]

@@

ggcorrplot(correlations$rho, hc.order = TRUE, outline.col = "white", type = "lower")
#p.mat <- cor_to_p(correlations$rho, 428, method = "polychoric")
#to show pvalue in the fig add p.mat=p.mat
```

@@

```{r EFA, include=FALSE}
fa.5F.1 <- fa(r=correlations$rho, nfactors = 5, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)
AA <- print(fa.5F.1, cut = .3, digits = 3, sort = TRUE)

@@

fa.5F.2 <- fa(r=correlations.red.5F.1$rho, nfactors = 5, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)
BB <- print(fa.5F.2, cut = .3, digits = 3, sort = TRUE)

@@

CC <- print(fa.5F.3, cut = .3, digits = 3, sort = TRUE)

@@

DD <- print(fa.5F.4, cut = .3, digits = 3, sort = TRUE)

@@

```{r include=FALSE}
fa.diagram(fa.5F.4,sort=TRUE,labels=NULL,flabels=NULL,cut=.3,
                 simple=TRUE,errors=F,
                 digits=1,e.size=.1,rsize=3,side=2,main= "Five Factor Solution Oblique",cex=NULL,color.lines=TRUE,
                 marg=c(.5,.5,1.5,.5),adj=2) 
```

```{r minres 5 factor supplimentary analyisis, eval=FALSE, include=FALSE}
fa.5F.min.1 <- fa(r=correlations$rho, nfactors = 5, fm= "minres",rotate ="varimax",
                residuals = TRUE, SMC = TRUE, n.obs =428)
EE <- print(fa.5F.min.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.5F.min.1 <- dplyr::select(LB ,
                                    - c(item20,item17,item3, item15,item40, item4,item11, item39,
                                        item18, item45, item29, item25, item8, item46 ))

correlations.red.5F.min.1 <- polychoric(reduced.model.5F.min.1,correct = 0)

fa.5F.min.2 <- fa(r=correlations.red.5F.min.1$rho, nfactors = 5, fm= "minres",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)
FF <- print(fa.5F.min.2, cut = .3, digits = 3, sort = TRUE)

reduced.model.5F.min.2 <- dplyr::select(LB,- c(item20,item17,item3, item15,item40, item4,item11, item39,
                                           item18, item45, item29, item25, item8, item46,
                                           item33, item12,item31, item37,
                                           item22, item38, item6, item48))

correlations.red.5F.min.2 <- polychoric(reduced.model.5F.min.2, correct = 0)

fa.5F.min.3 <- fa(r=correlations.red.5F.min.2$rho, nfactors = 5, fm= "minres",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)

GG <- print(fa.5F.min.3, cut = .3, digits = 3, sort = TRUE)

reduced.model.5F.min.3 <- dplyr::select(LB,- c(item20,item17,item3, item15,item40, item4,item11, item39,
                                               item18, item45, item29, item25, item8, item46,
                                               item33, item12,item31, item37,
                                               item22, item38, item6, item48,
                                               item9, item42, item23))
correlations.red.5F.min.3 <- polychoric(reduced.model.5F.min.3,correct = 0)

fa.5F.min.4 <- fa(r=correlations.red.5F.min.3$rho, nfactors = 5, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)

HH <- print(fa.5F.min.4, cut = .3, digits = 3, sort = TRUE)

```{r EFA 6 Factors Promax, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='hide', width=}
#LB.factor <- sapply(LB, as.factor)
#correlations <- polychoric(LB.factor, correct = 0)

fa.6F.1 <- fa(r=correlations$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)
II <- print(fa.6F.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.6F.1 <- dplyr::select(LB ,
                                    -c(item15, item17, item40, item4, item3, item46, item8,item29, item25 ))

correlations.red.6F.1 <- polychoric(reduced.model.6F.1,correct = 0)

fa.6F.2 <- fa(r=correlations.red.6F.1$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)
JJ <- print(fa.6F.2, cut = .3, digits = 3, sort = TRUE)

reduced.model.6F.2 <- dplyr::select(LB,-c(item15, item17, item40, item4, item3, item46, item8,item29, item25, item9, item33, item48, item16, item6 ))

correlations.red.6F.2 <- polychoric(reduced.model.6F.2,correct = 0)

fa.6F.3 <- fa(r=correlations.red.6F.2$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)

KK <- print(fa.6F.3, cut = .3, digits = 3, sort = TRUE)

reduced.model.6F.3 <- dplyr::select(LB,- -c(item15, item17, item40, item4, item3, item46, item8,item29, item25, item9, item33, item48, item16, item6, item11, item23, item18, item39))

correlations.red.6F.3 <- polychoric(reduced.model.6F.3,correct = 0)

fa.6F.4 <- fa(r=correlations.red.6F.3$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428, max.iter = 500)

LL <- print(fa.6F.4, cut = .3, digits = 3, sort = TRUE)

# Residuals
residual.6F =residuals(fa.6F.4 , diag = FALSE, na.rm =T)
#Count of numbers of residuals>.05
BigRRR= sum(residual.6F>abs(.05), na.rm =T)
print(BigRRR)

#Total number of off diagonal elements in the data matrix
totRRR = length(reduced.model.6F.3)*(length(reduced.model.6F.3)-1)/2

# Proportion of off-diagonal elements >.10
sumRRR <- sum(BigRRR/totRRR*100)

hist(residual.6F, col ="red", main = NULL, xlab = "Five factor oblique rotation: Residuals",ylim=c(0, 250))
abline(h =27, lwd = 2, lty = 2)
abline(v =.05,lwd = 2, lty = 2)
```

```{r factordetails, include=FALSE}
Fac.5 <- fa.5F.4 
#Commonality
min.com <- min(Fac.5$communality)
max.com <- max(Fac.5$communality)

# factor loadings

Factorloading <- cbind(Fac.5$loadings, Fac.5$communality)
colnames(Factorloading)= c("F1", "F2","F3", "F4", "F5", "Communalities")

min.loadings <- abs(min(Fac.5$loadings))
max.loadings <- max(Fac.5$loadings)

# Residuals
residual.5F =residuals(fa.5F.4, diag = FALSE, na.rm =T)
#Count of numbers of residuals>.05
BigRR= sum(residual.5F>abs(.05), na.rm =T)
print(BigRR)

#Total number of off diagonal elements in the data matrix
totRR = length(reduced.model.5F.3)*(length(reduced.model.5F.3)-1)/2

# Proportion of off-diagonal elements >.10
sumR <- sum(BigRR/totRR*100)

The initial five-factor solution with all 48 items showed the presence of cross-loading items (item 42, 16, & 1) and poor factor loading (<.30) items (item 20,3 ,15, 17, 40, 4, 11, 39, 18, 45, 29, 25, 8 , & 46). At first we discard the items with poor  factor loading and ran another EFA on the remaining 34 items. This iteration of EFA also appeared as a misfit in terms of poor factor loading (Item 12, 22, 38, 6) and cross-loading (items 23, 31, 37, 48) . Another two rounds of EFA were conducted with gradually identifying problematic items and discarding them from the model. Finally, a five-factor EFA solution with 23 items was accepted with low RMSR = 0.04, no loading smaller than .30 and no cross-loading greater than .30. The latent construct was also confirmed by using minimum residual  extraction method (see the supplementary). Table\@ref(tab:tabEFA5) displays the structural coefficients and commonality of the items. The absolute value of the structural coefficients ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. However, the histogram of the absolute values of non-redundant residual-correlations (Fig\@ref(fig:Residuals) showed `r apa(sumR,2, T)`% correlations greater than the absolute value of .05, indicating under-factoring. [@desjardinsHandbookEducationalMeasurement2018a]. Subsequently, we fitted a six-factor solution. However, in the six factor solution a factor emerged with only two salient variable loading thus disqualifying the six-factor solution.

In the five

```{r TabEFA5, results='asis'}

apa_table(unclass(Factorloading))
```

```{r Residuals, out.height= "50%",  fig.align='center', fig.cap=' Histogram of residulas:  five-factor solution,  out.width="50%",  warning=FALSE'}
hist(residual.5F, col ="red", main = NULL, xlab = "Five-factor solution", ylim=c(0, 250))
abline(h =47, lwd = 2, lty = 2)
abline(v =.05,lwd = 2, lty = 2)

### 2021-08-31 manuscript.tex

We used R (version 4.1.0), including several R-packages for our analyses. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data we used polychoric correlation matrix (Desjardins \& Bulut, 2018) for the EFA. We employed principal axis (pa) a factor extraction method with varimax rotation. PA is apparently robust to the normality assumption violations (Watkins, 2020). The obtained latent structure was confirmed by weighted least squares (WLS) extraction method as well. We used a combination factor indentification method including scree plot(Cattell, 1966), Horn's parallel analysis (Horn, 1965), minimum average partials method(Velicer, 1976), and hull method (Lorenzo-Seva, Timmerman, \& Kiers, 2011) to identify factor numbers. Additionally, to identify the simple structure we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading \textless0.3 (iii) no items with cross-loading greater than .3 across factors (Bandalos \& Finney, 2018; Child, 2006; Mulaik, 2009; Watkins, 2020)

@@

Scree plot ( Fig@ref(fig:fac.id)) suggested a six-factor solution.Horn's parallel analysis (Horn, 1965), like the Monte Carlo study, draws several sets of random data with the same number of participants as the original data set and compares the mean eigenvalues among the simulated and original data sets to retain optimal factors.This extraction method also supported a five-factor model. In our data set parallel analysis with 500 iterations indicated six-factor solution. However, In MAP method (Velicer, 1976) and Hull method (Lorenzo-Seva, Timmerman, \& Kiers, 2011) suggested a five-factor solution. As a result, we tested both five factor and six factor solutions. The histogram of the absolute values of non-redundant residual-correlations the proportion of non-redundant residual correlations greater than the absolute value of .05 should be small

@@

@@

The initial five-factor solution with all 48 items showed the presence of cross-loading items (item 42, 16, \& 1) and poor factor loading (\textless.30) items (item 20,3 ,15, 17, 40, 4, 11, 39, 18, 45, 29, 25, 8 , \& 46). At first we discard the items with poor factor loading and ran another EFA on the remaining 34 items. This iteration of EFA also appeared as a misfit in terms of poor factor loading (Item 12, 22, 38, 6) and cross-loading (items 23, 31, 37, 48) . Another two rounds of EFA were conducted with gradually identifying problematic items and discarding them from the model. Finally, a five-factor EFA solution with 23 items was accepted with RMSR = 0.04, no loading smaller than .30 and no cross-loading greater than .30.

>>>>

We used R (version 4.1.0), including several R-packages for our analyses. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data we used polychoric correlation matrix (Desjardins \& Bulut, 2018) for the EFA. We employed principal axis (pa) a factor extraction method with varimax rotation. PA is apparently robust to the normality assumption violations (Watkins, 2020). The obtained latent structure was confirmed by minimum residuals extraction method as well. We used a combination factor indentification method including scree plot(Cattell, 1966), Horn's parallel analysis (Horn, 1965), minimum average partials method(Velicer, 1976), and hull method (Lorenzo-Seva, Timmerman, \& Kiers, 2011) to identify factor numbers. Additionally, to identify the simple structure we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading \textless0.3 (iii) no items with cross-loading greater than .3 across factors (Bandalos \& Finney, 2018; Child, 2006; Mulaik, 2009; Watkins, 2020)

@@

Scree plot ( Fig@ref(fig:fac.id)) suggested a six-factor solution.Horn's parallel analysis (Horn, 1965), like the Monte Carlo study, draws several sets of random data with the same number of participants as the original data set and compares the mean eigenvalues among the simulated and original data sets to retain optimal factors.This extraction method also supported a five-factor model. In our data set parallel analysis with 500 iterations indicated six-factor solution. However, In MAP method (Velicer, 1976) and Hull method (Lorenzo-Seva, Timmerman, \& Kiers, 2011) suggested a five-factor solution. As a result, we tested both five factor and six factor solutions.

@@

The initial five-factor solution with all 48 items showed the presence of cross-loading items (item 42, 16, \& 1) and poor factor loading (\textless.30) items (item 20,3 ,15, 17, 40, 4, 11, 39, 18, 45, 29, 25, 8 , \& 46). At first we discard the items with poor factor loading and ran another EFA on the remaining 34 items. This iteration of EFA also appeared as a misfit in terms of poor factor loading (Item 12, 22, 38, 6) and cross-loading (items 23, 31, 37, 48) . Another two rounds of EFA were conducted with gradually identifying problematic items and discarding them from the model. Finally, a five-factor EFA solution with 23 items was accepted with low RMSR = 0.04, no loading smaller than .30 and no cross-loading greater than .30. The latent construct was also confirmed by using minimum residual extraction method (see the supplementary). Table\ref{tab:tabEFA5} displays the structural coefficients and commonality of the items. The absolute value of the structural coefficients ranged from .47 to .99 indicating strong coefficients. The commonalities ranged between .10 to .99. However, the histogram of the absolute values of non-redundant residual-correlations showed 26.09\% correlations greater than the absolute value of .05, indicating under factoring. (\textbf{desjardinsHandbookEducationalMeasurement2018a?}).

@@
